Certified Knowledge Compilation \\ with Application to Verified Model Counting

### 2023-03-15 doc/mc-data.tex
Did not meet criteria

\section*{Methodology}
\begin{enumerate}
   \item Retrieved 200 public model counting benchmarks from 2022 Model Counting Competition website
     \begin{center}
       \url{https://mccompetition.org/2022/mc_description.html}
     \end{center}
     \begin{itemize}
     \item 100 public instances from Track 1 (unweighted model counting)
     \item 100 public instances from Track 2 (weighted model counting)
     \end{itemize}
   \item Removed duplicates
     \begin{itemize}
    \item  20 instances were duplicated in the two tracks
     \item Reduces set of benchmarks to 180
     \end{itemize}
   \item Ran D4 on 180 instances with time limit of 4000 seconds
     \begin{itemize}
     \item Completed for 124 instances
     \end{itemize}
   \item Generated POG representations from dDNNF graphs
     \begin{itemize}
     \item Simplifications via constant propagation
       \begin{itemize}
       \item Number of POG nodes around $0{.}90\times$ number of dDNNF nodes
       \end{itemize}
     \item Used heuristics to identify root node, and then put in topological order with root node at end.
     \item Measure complexity by number of defining clauses $D$
       \begin{itemize}
       \item Product or sum node with degree $k$ requires $k+1$ clauses
       \item $D$ also equals sum of number of edges and number of nodes
       \item For 124 instances, Min = 304, Median = 848,784, Mean = 93,107,937, Max = 2,761,457,765
       \end{itemize}
     \end{itemize}
   \item Ran program {\sc d2p} to generate CRAT file for maximum of 10,000 seconds
     \begin{itemize}
     \item Completed for 108 instances
     \item Exceeded clause limit for 1 instance (Our programs represent a clause ID as a 32-bit signed value)
     \item Ran out of memory for 1 instance
     \item Timed out for 14 instances
     \end{itemize}
   \item For completed CRAT files, ran {\sc crat-check} CRAT checker
     \begin{itemize}
     \item Tests all conditions for correct CRAT
     \item Most time spent performing RUP checks for clause addition and deletion
     \item Completed successfully for all 108 instances
     \item Time generally faster than {\sc d2p}. (min = 0.005, max = 1.4, harmonic mean = 0.10)
     \end{itemize}
\end{enumerate}

@@

\begin{figure}
\centering{%
\begin{tikzpicture}[scale = 0.90]
  \begin{axis}[mark options={scale=0.55},grid=both, grid style={black!10}, ymode=log,
      legend style={at={(0.22,0.98)}},
      legend cell align={left},
                              x post scale=1.8, y post scale=2.2,
                              xmode=log,xmin=0.01,xmax=1500,
                              xtick={0.01, 0.1,1.0,10,100,1000}, xticklabels={0.01, 0.1, 1.0, 10, 100,{1,000}},
                              ymin=0.01, ymax=1900,
                              ytick={0.01, 0.1,1.0,10,100,1000}, yticklabels={0.01, 0.1, 1.0, 10, 100, {1,000}},
                              xlabel={Structured generation + checking (seconds)}, ylabel={Monolithic generation + checking (seconds)},
            ]
    \input{data-formatted/time-structured+monolithic}
    \input{data-formatted/time-structured+monolithic_timeout}
    \addplot[mark=none, color=lightblue] coordinates{(1.0,0.01) (10000.0,100.0)};
    \addplot[mark=none, color=lightblue] coordinates{(0.1,0.01) (10000.0,1000.0)};
    \addplot[mark=none, color=lightblue] coordinates{(0.01,0.01) (10000.0,10000.0)};
    \addplot[mark=none, color=lightblue] coordinates{(0.01,0.1) (10000,100000)};
    \addplot[mark=none, color=lightblue] coordinates{(0.01,1.0) (1000,100000)};
    \addplot[color=black,dashed] coordinates{(0.01,1500) (1500,1500)};
    \node[left] at (axis cs: 5.0,0.06) {$0{.}01\times$};
    \node[left] at (axis cs: 0.5,0.06) {$0{.}1\times$};
    \node[left] at (axis cs: 0.05,0.06) {$1\times$};
    \node[left] at (axis cs: 0.05,0.6) {$10\times$};
    \node[left] at (axis cs: 0.05,6.0) {$100\times$};

@@

@@

For comparison, we also ran the MICE tool chain for certifying model
counters~\cite{fichte:sat:2022}.  Figure~\ref{fig:mice} compares the
runtimes of our toolchain versus theirs.  The X-axis shows the sum of
the elapsed times for our generator, checker, and counter, while the
Y-axis shows the sum of the elapsed times for \textsc{nnf2trace},
generating a file in their trace format from a d-DNNF graph generated
by D4, and for \textsc{sharptrace}, a checker for the trace format.
The plot shows data points for those benchmarks for at least one of
the two tools completed in less than 1000 seconds.

@@

Overall, we can summarize the results as:
\begin{itemize}
\item 85 total data points
\item 64 were completed by both programs in under 1000 seconds
\item Of these, the MICE toolchain was faster for 20, and the CRAT toolchain was faster for 45
\item 5 completed with CRAT under 1000 seconds but exceeded that threshold for MICE
\item 1 completed with MICE under 1000 seconds but exceeded that threshold for CRAT
\item 15 completed with CRAT but generated errors for MICE.
\end{itemize}

>>>>

\section*{Methodology}
\begin{enumerate}
   \item Retrieved 200 public model counting benchmarks from 2022 Model Counting Competition website
     \begin{center}
       \url{https://mccompetition.org/2022/mc_description.html}
     \end{center}
     \begin{itemize}
     \item 100 public instances from Track 1 (unweighted model counting)
     \item 100 public instances from Track 2 (weighted model counting)
     \end{itemize}
   \item Removed duplicates
     \begin{itemize}
    \item  20 instances were duplicated in the two tracks
     \item Reduces set of benchmarks to 180
     \end{itemize}
   \item Ran D4 on 180 instances with time limit of 4000 seconds
     \begin{itemize}
     \item Completed for 124 instances
     \end{itemize}
   \item Generated POG representations from dDNNF graphs
     \begin{itemize}
     \item Simplifications via constant propagation
       \begin{itemize}
       \item Number of POG nodes around $0{.}90\times$ number of dDNNF nodes
       \end{itemize}
     \item Used heuristics to identify root node, and then put in topological order with root node at end.
     \item Measure complexity by number of defining clauses $D$
       \begin{itemize}
       \item Product or sum node with degree $k$ requires $k+1$ clauses
       \item $D$ also equals sum of number of edges and number of nodes
       \item For 124 instances, Min = 304, Median = 848,784, Mean = 93,107,937, Max = 2,761,457,765
       \end{itemize}
     \end{itemize}
   \item Ran program {\sc d2p} to generate CRAT file for maximum of 10,000 seconds
     \begin{itemize}
     \item Completed for 108 instances
     \item Exceeded clause limit for 1 instance (Our programs represent a clause ID as a 32-bit signed value)
     \item Ran out of memory for 1 instance
     \item Timed out for 14 instances
     \item Was able to do one-sided verification for 9 of these.
     \end{itemize}
   \item For completed CRAT files, ran {\sc crat-check} CRAT checker
     \begin{itemize}
     \item Tests all conditions for correct CRAT
     \item Most time spent performing RUP checks for clause addition and deletion
     \item Completed successfully for all 108 instances
     \item Time generally faster than {\sc d2p}. (min = 0.005, max = 1.4, harmonic mean = 0.10)
     \end{itemize}
\end{enumerate}

@@

\begin{figure}
\centering{%
\begin{tikzpicture}[scale = 0.90]
  \begin{axis}[mark options={scale=0.55},grid=both, grid style={black!10}, ymode=log,
      legend style={at={(0.22,0.98)}},
      legend cell align={left},
                              x post scale=1.8, y post scale=2.2,
                              xmode=log,xmin=0.01,xmax=1900,
                              xtick={0.01, 0.1,1.0,10,100,1000}, xticklabels={0.01, 0.1, 1.0, 10, 100,{1,000}},
                              ymin=0.01, ymax=1900,
                              ytick={0.01, 0.1,1.0,10,100,1000}, yticklabels={0.01, 0.1, 1.0, 10, 100, {1,000}},
                              xlabel={Structured generation + checking (seconds)}, ylabel={Monolithic generation + checking (seconds)},
            ]
    \input{data-formatted/time-structured+monolithic}
    \input{data-formatted/time-structured+monolithic_timeout}
    \addplot[mark=none, color=lightblue] coordinates{(1.0,0.01) (10000.0,100.0)};
    \addplot[mark=none, color=lightblue] coordinates{(0.1,0.01) (10000.0,1000.0)};
    \addplot[mark=none, color=lightblue] coordinates{(0.01,0.01) (10000.0,10000.0)};
    \addplot[mark=none, color=lightblue] coordinates{(0.01,0.1) (10000,100000)};
    \addplot[mark=none, color=lightblue] coordinates{(0.01,1.0) (1000,100000)};
    \addplot[color=black,dashed] coordinates{(0.01,1500) (1500,1500)};
    \addplot[color=black,dashed] coordinates{(1500,0.01) (1500,1500)};
    \node[left] at (axis cs: 5.0,0.06) {$0{.}01\times$};
    \node[left] at (axis cs: 0.5,0.06) {$0{.}1\times$};
    \node[left] at (axis cs: 0.05,0.06) {$1\times$};
    \node[left] at (axis cs: 0.05,0.6) {$10\times$};
    \node[left] at (axis cs: 0.05,6.0) {$100\times$};

@@

Ran all benchmarks for which generation + checking could complete in 1000 seconds.  Total = 83.  All completed in less than 1000 seconds with structured.
Compare both runtimes and total number of clauses for structured vs. monolithic:
\begin{itemize}
\item Structured complete 83.
\item Monolithic completed  69, timed out on 14.
\item Of the 69 completed, 23 were faster with monolithic, 36 were slower (min ratio = 0.015, max ratio = 9.66, hmean = 0.40)
\item Of the 69 completed, 59 had fewer total clauses with monolithic, 10 had more (min ratio = 0.07, max ratio = 1.09, hmean = 0.49)
\end{itemize}

@@

For comparison, we also ran the MICE tool chain for certifying model
counters~\cite{fichte:sat:2022}.  Figure~\ref{fig:mice} compares the
runtimes of our toolchain versus theirs.  The X-axis shows the sum of
the elapsed times for our generator, checker, and counter (since the MICE checker also provides a count), while the
Y-axis shows the sum of the elapsed times for \textsc{nnf2trace},
generating a file in their trace format from a d-DNNF graph generated
by D4, and for \textsc{sharptrace}, a checker for the trace format.
The plot shows data points for those benchmarks for at least one of
the two tools completed in less than 1000 seconds.

@@

Overall, we can summarize the results as:
\begin{itemize}
\item 84 total data points
\item 63 were completed by both programs in under 1000 seconds
\item Of those 20 were faster with MICE, 43 with CRAT
\item 6 completed with CRAT under 1000 seconds but exceeded that threshold for MICE
\item 1 completed with MICE under 1000 seconds but exceeded that threshold for CRAT
\item 14 completed with CRAT but generated errors for MICE.
\end{itemize}
